---
title: "DS2 HW3"
author: "John Neeld"
format: docx
execute:
  message: false
  warning: false
  echo: true
editor: 
  markdown: 
    wrap: sentence
---

The `babynames` dataset in the `babynames` packages contains all births in the U.S. from 1880 to 2015 in which there were at least 5 births of that *sex-name* dyad in that year.
(A "dyad" is a combination of two variables, "triad" is three, etc.)

Create a dataset called `babynames_total` that totals `births` across all years.
(Note: The case in the new dataset should be `sex` and `name`, which means that each of these dyads should appear only once in the dataset.)

Use this dataset to calculate what is asked below:

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(babynames)
library(tinytex)
```

## 1) 10 most popular names

```{r}
babynames_total <- babynames |> 
  select(sex,name, n) |> 
  group_by(sex,name) |> 
  summarise(births = sum(n),
            .groups = "drop")

babynames_total |> 
  arrange(desc(births)) |> 
  head(10)
```

James, John, Robert, Michael, Mary, William, David, Joseph, Richard, and Charles are the 10 most popular names.

## 2) Least popular names

```{r}
babynames_total |> 
  arrange(births)

babynames_total |> 
  count(births)
```

a.  What is the smallest value of `births` in the dataset?

    *JOE: Fixed a small typo*

    a.  **The smallest value of births in the dataset is 5.**

b.  How many *sex-name* have that many births?

    a.  **There are 15477 sex-name that have 5 births.**

c.  Based on the way the data were assembled, explain why there are so many.

    a.  **I believe that it was because people accidentally spelled the name, they intended to write, wrong on the birth certificate.**

## 3) 10 most popular [boys']{.underline} names that end in a vowel

```{r}
bnt_vowel <- babynames_total |> 
  mutate(last_char = str_sub(name, -1, -1),
         vowelend = last_char %in% c("a", "e", "i", "o", "u"))
  
bnt_vowel |> 
  filter(sex == "M") |> 
  filter(vowelend == TRUE) |> 
  arrange(desc(births)) |> 
  head(10)
```

George, Joshua, Jose, Kyle, Lawrence, Joe, Willie, Jesse, Bruce, and Eugene are the 10 most popular boys' names that end in a vowel.

*JOE: I like that you included the important output from the code as text that doesn't rely on the code itself. Redundancy is often a good thing when programming.*

## 4) 10 most popular [girls']{.underline} names with joe/jo suffix

Find the 10 most popular girls' names that end in "joe" or "jo", excluding the names "Joe", and "Jo".
For this, problem, you should include all spellings, whether or not they are capitalized (so, for example, "Mary Jo" counts as ending in "jo")

```{r}
babynames_total |> 
  filter(sex == "F") |> 
  mutate(jos = name |> str_sub(-3) |> str_count("joe"),
         jos = name |> str_sub(-2) |> str_count("jo")) |> 
  filter(jos > 0) |> 
  arrange(desc(births)) |> 
  head(10)

```

Mary Jo, Billie Jo, Mari Jo, Bobbi Jo, Bobbie Jo, Betty Jo, Amy Jo, Samantha Jo, Mar Jo, and Kelly Jo are the 10 most popular girls' names with a joe/jo suffix.

## 5) Allnameslikethis

Did we need to worry about case in the last problem?
To answer this, verify the following:

*JOE: Changed the alphanumeric markers from "a, a, a"* *to "a, b, c"*

a.  There are no spaces in any names

```{r}
babynames_total |> 
  as.data.frame() |> 
  mutate(sp = str_detect(name, " ")) |> 
  count(sp)
```

b.  The first letter is always capitalized

```{r}
babynames_total |> 
  mutate(cap = name |> str_sub(1,1) |> str_count("[a-z]")) |> 
  count(cap)
```

c.  No other letters are capitalized

```{r}
babynames_total |> 
  # JOE: Shorthand for listing the whole alphabet is 'A-Z', saves time; also did it in b's block above
  mutate(cap = name |> str_sub(2) |> str_count("[A-Z]")) |> 
  count(cap)
```

#### The problems below are optional, since they are not working with text, but I thought they provided an interesting challenge.

## 6) Closest to the average (optional)

Calculate the average number of births by *sex-name*.
Which *sex-name* is closest to that number?
Do this algorithmically in one pipeline, meaning no saving any objects and no typing in the average yourself.

```{r}

```

## 7) 10 most common cross-sex names (optional)

Show the most common names that are given both to males and to females.
You will have to make a decision about how to quantify this - I recommend not looking just at total births (male + female).

```{r}

```
